---
output: github_document
editor_options: 
  markdown: 
    wrap: 80
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

set.seed(1234)
```

# mifa - Multiple Imputation for Exploratory Factor Analysis

<!-- badges: start -->

[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![R build status](https://github.com/teebusch/mifa/workflows/R-CMD-check/badge.svg)](https://github.com/teebusch/mifa/actions)
<!-- badges: end -->

This package uses multiple imputation to estimate the covariance matrix of
incomplete data. An exploratory factor analysis then can be applied on this
estimated covariance matrix. It also provides Fieller and bootstrap confidence
intervals for the proportion of explained variance using different number of
factors.

#### For more information, see:

Nassiri, V., Lovik, A., Molenberghs, G. *et al.* On using multiple imputation
for exploratory factor analysis of incomplete data. *Behav Res* 50, 501–517
(2018). <https://doi.org/10.3758/s13428-017-1013-4>

## Installation

You can install the the release version of mifa from
[GitHub](https://github.com/) with:

```r
# install.packages("devtools")
devtools::install_github("vahidnassiri/mifa")
```

## Usage

### Generate Data Set

Genarate a data set with 5% of the data missing completely at random (MCAR)

```{r generate-data, message=FALSE}

library(eigeninv)

e.vals <- c(50, 48, 45, 25, 20, 10, 5, 5, 1, 1, 0.5, 0.5, 0.5, 0.1, 0.1) # eigenvalues
P      <- length(e.vals)  # number of items
N      <- 100             # sample size

# generate a covariance matrix with the eigenvalues in e.vals
cov.mat  <- eigeninv::eiginv(evals = e.vals, symmetric = TRUE) 
chol.cov <- t(chol(cov.mat)) # Cholesky decomposition of the covariance matrix

# Generate centered independent normal data
data.ini1     <- matrix(rnorm(N * P), N, P)
mean.data.ini <- apply(data.ini1, 2, mean)
data.ini      <- t(t(data.ini1) - mean.data.ini)

# generate multivariate normal data with the given covariance matrix.
data <- matrix(0, N, P)
for (i in 1:N) {
  data[i, ] <- chol.cov %*% data.ini[i, ]
}

# Create 5-percent missing data with missing completely at random
data.miss <- data
mcar.n.miss <- 0.05

for (i in 1:P) {
  for (j in 1:N) {
    rand.u <- runif(1)
    if (rand.u <= mcar.n.miss) {
      data.miss[j, i] <- NA
    }
  }
}
```

### Run mifa

Run mifa:

```{r run-mifa, messages=FALSE}

library(mifa)

result.mi <- mifa.cov(data.miss,
  n.factor = 1:10, M = 10, maxit.mi = 5, method.mi = "pmm",
  alpha = 0.05, rep.boot = 50, ci = TRUE
)

summary(result.mi)
```
The Fieller confidence intervals indicate that 4 factors are enough to explain 
around 80% of the variance.

```{r}
round(result.mi$ci.mice.fieller, 2)
```
The estimated covariance matrix based on imputed data is in `result.mi$cov.mice`, 
and we can use it to perform exploratory factor analysis:

```{r}
library(psych)
fit <- fa(r = result.mi$cov.mice, n.obs = N, nfactors = 4, rotate = "varimax")

fit

fa.diagram(fit)
```
`mifa.cov()` provides the confidence intervals when `ci == TRUE`. 
CIs can also be calculated separately if desired:

```{r run-mifa-ci}

# get CIs using Fieller's method
result.mi <- mifa.cov(data.miss,
  n.factor = 1:10, M = 10, maxit.mi = 5, method.mi = "pmm",
  alpha = 0.05, rep.boot = 500, ci = FALSE
)

ci.mifa.fieller(result.mi$cov.mice.imp, 1:10, 0.05, N = 100)


# Get CIs using bootstrapping
ci.mifa.bootstrap(data.miss, 
  n.factor = 1:10, rep.boot = 100, maxit.mi = 5, method.mi = "pmm", 
  alpha = 0.05
)

```
