---
output: github_document
editor_options: 
  markdown: 
    wrap: 80
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mifa - Multiple Imputation for Exploratory Factor Analysis

<!-- badges: start -->

<!-- badges: end -->

This package uses multiple imputation to estimate the covariance matrix of
incomplete data. An exploratory factor analysis then can be applied on this
estimated covariance matrix. It also provides Fieller and bootstrap confidence
intervals for the proportion of explained variance using different number of
factors.

#### For more information, see:

Nassiri, V., Lovik, A., Molenberghs, G. *et al.* On using multiple imputation
for exploratory factor analysis of incomplete data. *Behav Res* 50, 501–517
(2018). <https://doi.org/10.3758/s13428-017-1013-4>

## Installation

You can install the the release version of mifa from
[GitHub](https://github.com/) with:

```r
# install.packages("devtools")
devtools::install_github("vahidnassiri/mifa")
```

## Usage

```{r example}
library(mifa)
```
